---
title: "R_Part_3"
author: "Thilina"
date: "8/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

INSTALLATION
you can either install the whole "tidyverse" package or "ggplo2" package to create 
graphics
Then you have to load the package
     1.library(tidyverse)
     2.library(ggplot2)
```{r}
library(ggplot2)
```
Basic Syntax
   ggplot(data=<Data>,aes(x=  ,y=  ))+geo_function()
  
1.Scater Plot
```{r}
data("msleep")
head(msleep,8)
```
How do we plot 'bodywt' againts 'sleeep_total' in a scatter plot using 'ggplot2' 
package in Rstudio?
```{r}
#scattter plot

scatterplot_1<-ggplot(data=msleep,aes(x=bodywt,y=sleep_total))+geom_point()
scatterplot_1
```
#colour
```{r}

scatterplot_1<-ggplot(data=msleep,aes(x=bodywt,y=sleep_total))+geom_point(col="blue")
scatterplot_1
```
2.Histogram
```{r}
set.seed(1234)
weight_data<-data.frame(sex=factor(rep(c("F","M"),each=200)),weight=round(c(rnorm(200,mean=55,sd=5),rnorm(200,mean=75,sd=3))))
head(weight_data)
```
How do we plot 'aweight' in histogram using 'ggplot2' 
```{r}
#Hiistogram
ggplot(data=weight_data,aes(x=weight))+geom_histogram()
```
```{r}
#Hiistogram 
#bin width
#color of margins
#margin line type
ggplot(data=weight_data,aes(x=weight))+geom_histogram(binwidth =1,fill="green",color="black",linetype="dashed")
```
##for linetype 
```{r}
?linetype
```
 
3.Box Plot
```{r}
data("airquality")
airquality

```
```{r}
str(airquality)
```

How to make box plots for the 'ozone'  readings (for each month separately) using 'ggplot2'
```{r}
#convert Monthh into a labeled factor
airquality$Month=factor(airquality$Month,labels=c("May","Jun","Jul","Aug","Sep"))
str(airquality)
```

```{r}
#basic box plot
b1<-ggplot(data=airquality,aes(x=Month,y=Ozone))+geom_boxplot(fill="green",col="red",outlier.colour = "blue"
,outlier.shape =5 )+xlab("Month")+ylab("Mean Ozone Concenration")+ggtitle("Box Plot")
b1
```
```{r}
#Defferent colurs for box plot
#possition of legend
b1<-ggplot(data=airquality,aes(x=Month,y=Ozone,fill=Month))+geom_boxplot(col="red",outlier.colour = "blue"
,outlier.shape =5 )+xlab("Month")+ylab("Mean Ozone Concenration")+ggtitle("Box Plot")+scale_fill_manual(values=c("orange","red","blue","pink","green"))+theme(legend.position = "top")
b1
```

4.Bar chart
```{r}
data(mtcars)
head(mtcars,10)
str(mtcars)
```
How to create bar chart for 'cyl' 
```{r}
#Basic bar chart
ggplot(mtcars,aes(x=as.factor(cyl)))+geom_bar(fill=heat.colors(3),col="black",width=0.5)
```
```{r}
#Basic bar chart colour palat
ggplot(mtcars,aes(x=as.factor(cyl),fill=as.factor(cyl)))+geom_bar(color="black",width=0.5)+scale_fill_brewer(palette="Dark2")
```

geom_bar can be used for these occasions also
```{r}
data<-data.frame(name=c("A","B","C","D","E"),value=c(7,15,6,14,9))
data
```
```{r}
ggplot(data,aes(x=name,y=value))+geom_bar(stat="identity")
```

5.Pie Chart
```{r}
#Create the data frame
data<-data.frame(subject=c("Physics","English","Mathematics","Biology","Chemistry"),
count=c(75,45,62,95,34))
data
```

How to plot above data in a pie chart
#stat bar char
```{r}
ggplot(data,aes(x="",y=count,fill=subject))+geom_bar(stat="identity",width=1)
```
```{r}
#Basic pie chart
ggplot(data,aes(x="",y=count,fill=subject))+geom_bar(stat="identity",width=1)+coord_polar("y",start=0)
```

# slide presentation of pie chart


