---
title: "R_Code_2"
author: "Thilina"
date: "8/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

ITRODUCTION
A function.in a programming environment,is a set of instructions organized togeter to perform a specific task.
programer builds a function to avoid repeationg the same task,or reduce complesity.

There are two types of R functions
   *Built-in R functions
   *User-Defined functions
   
#1)Built-in R functions
The functions which are already created or defined in programming framework.
We can see 3 Groups of functions in action.
 i)Genarla function
 ii)Math functions
 iii)statistical function
 
i)Genaral Functions

Functions We used to perform genaral actions
EX: cbind(),rbind(),sort(),length(),ect..

```{r}
X<-c(8,4,6,7,2,9,12,3,15,11)
sort(X)
length(X)
```
ii)Math functions

Functions we used to perform mathametical operations
EX:  abs(),sqrt(),fractorial()
```{r}
abs(-8)
sqrt(4)
factorial(5)
```
iii)Statistical Functions

Functions we used to perform statistical analysis
EX: mean(),var(),sd(),median()
```{r}
x<-c(8,4,6,7,2,9,12,3,15,11)
mean(x)
var(x)
sd(x)
median(x)
```

#2).User Defined Function

An R function is created by using the keyword function. The basic syntax of an R function definition is as follows 
function_name <- function(arg_1, arg_2, ...) {
   Function body 
}

Devlop a function to add two numbers
```{r}
function_1<-function(x,y){
  return(x+y)
}
```
Calling the function
  Function_name(input value for argument1,input value for agument2,..)
```{r}
function_1(2,1)
```

Functions without arguments
```{r}
function_2<-function(){
  for(i in 1:5){
    print(factorial(i))
  }
}
```
```{r}
function_2()
```

Function with arguments
```{r}
function_3<-function(a,b=4){
  return(a*b)
}
```


```{r}
function_3(8)
```

SPECIAL NOTE ON FUNCTIONS
*print the aruments and corresponding default values for a function
     syntax:args(Function_name)
```{r}
args(function_3)
```

##print the content of a function by typing only function without any brackets
   syntex:Function_name
```{r}
function_3
```

##Edit a function
 syntex:Function_name<-edit(Function_name)
 
```{r}
function_4<-edit(function_3)
function_4
```

#Control Structures
control structures allow you to put some “logic” into your R code, rather than just always executing the same R code every time

1)IF CONDITION
The take is carried out only if the test expression is returmed as TRUE.If the test 
expression is false,nothing happens.

syntex: if(test_expression){
          statement
        }
EX:
```{r}
values<-1:10
y<-sample(values,1)
 if (y<=5){
   print(paste(y,"Is less thgan or equal 5"))
 }
```

IF-ELSE CONDITION
The if-else combination is probably the most commonly used control structure in R (or perhaps any language). This structure allows you to test a condition and act on it depending on whether it’s true or false

Syntex:if(<condition>) {
        ## do something
        } 
        else {
        ## do something else
        }
Ex:
```{r}
val1=9#Creating first variable
val2=5 #Creating second variable
if (val1<val2){
  print("value 2 is greater than value 1")
}else{
    print("value 1 is greater than value 2")
  }
```
There is a vec tor equivalent from of the ifelse statement
 syntex: 
           ifelse(test_expresson,x,y)

This expression returns a vector.This returned vector has element from X if the 
corresponding value of test_expression is TRUE or from Y,if the corresponding 
value of test_expression is FALSE
```{r}
a<-c(5,7,2,9,6)

ifelse(a %% 2==0,"even","odd")
```

IF-ELSE STATEMENT WITH MULTIPLE CONDITIONS(IF-ELSE LADDER)
when we want to perform various actions based on verious conditions,we can use if-else statement more thane one
Sintax:
       if(<condition1>) {
        ## do something
        } else if(<condition2>)  {
        ## do something different
        } else {
        ## do something different
        }
        
EX: 
```{r}
sing<-function(X){
  if(X==0){
    result<-"Zero"
  }else if(X<0){
    result<-"Negative"
  }else {
    result<-"Positive"
  }
  return(result)
}
sing(0)
sing(-8)
sing(8)
```

At times,we may face a situation in which we nedda to execute a statement or a group of statements
(block of codes) multple times.
This concpt is known as looping and the programming construct which allows the repetition 
is known as the loop.
A loop statement allows us to execute a statement or group of statements
multiple times.

So here we discuss 3 types of loops
  *for
  *while
  *repeat
When we want to change execution of loops from normal sequence,we use
  *next
  *break
  
2)For Loop
For loops are most commonly used for iterating over the elements of an object (list, vector, etc.)
 Syntex: 
 for (i in 1:n){
   statement
 }
 
EX:

```{r}
for(i in 1:10){
  print(i+5)
}
```

3)While Loop
A while loop is used to repeat an statement untill a specific condition is met
syntex:
      while(condition){
      statements
      }
EX:
```{r}
students<-c("kamal","Nizar","Sachini")
count=3
    while(count<10){
      print(students)
      count=count+1
      }
```

4)Repeat Loop
A repeat loop is used to iterate over a block of code multiple times.In this loop,
no condition checked to exit the loop.

To stop repeating the loop,we must put a condition explicity inside the body of the loop
and use the break statement to exit the loop.

If you failed to do so,it will create an infinite loop.
Syntax:
       repeat{
       statement
       }
      
##Brak Statement     
A break ststement is used inside a loop(repeat,for,while)to stop the iteration and flow the control outside of the loop.

Syntax of using break statement within an if statement:
   if(Test_expression){
    break
   }
EX:
```{r}
z<-0
  repeat{
    print(z)
    if(z==6){
      break
    }
    z<-z+2
  }
   
```
##Next Statement
a next statement is used to skip the current iteration in the loop and move to the next 
iteration without exiting from the loop itself.

Syntax of using next statement within an if statement:
 if(test_condition){
   next
 }
Ex:
```{r}
x<-1
 while(x<5){
   x<-x+1
   if (x==3){
     next
   }
   print(x)
 }
```

